Housing Market Institutions Drive Race and Ethnicity Differences in Energy Consumption

### 2015-03-25 ConferencePaper/DraftEdwardsWoods.Rnw
Added a bunch of text about the metrics but have a lot commented out traceing a pdf compile problem.  Working on solutions.

\cite{RBase}

@@

<<tabRaceVStruct,echo=false,results=tex>>=
library(xtable)
xtable(table(RECS$EthRace,RECS$TYPEHUQ), caption="Count of Observations by Race and Structure Type", label="tab:RaceVStruct")
@

<<echo=false,results=tex>>=
library(xtable)
xtable(table(RECS$REGIONC,RECS$TYPEHUQ), caption="Count of Observations by Region and Structure Type", label="tab:RegionVStruct")
@

\begin{figure}
\begin{center}
\caption{Annual kWh by Rent/Own and Race}
\label{fig:kWhbyOwnRace}
<<echo=false,fig=TRUE>>=
library(ggplot2)
qplot(KWH, data = RECS, xlim=c(0,30000),facets =EthRace~KOWNRENT, geom = "density" )
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Annual kWh by Income}
\label{fig:kWhbyIncome}
<<echo=false,fig=TRUE>>=
library(ggplot2)
qplot(Income,KWH, data = RECS, xlim=c(0,30000),facets =EthRace~KOWNRENT )+geom_boxplot()
@
\end{center}
\end{figure}

<<echo=false,fig=TRUE>>=
qplot(NHSLDMEM,data=RECS,facets =~EthRace, geom = "density",title="Household Size by Race and Ethnicity")+xlab("Household Size")
@

@@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$UR),caption="Race by Urban/Rural", label="tab:UrbRural")
@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$ESWWAC),caption="EnergyStar Wall AC by Race and Ethnicity", label="tab:EstarWall")
@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$UR),caption="Race by Urban/Rural", label="tab:UrbRural")
@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$FUELH2O),caption="Water Service by Race and Ethnicity", label="tab:Water")
@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$FUELHEAT),caption="Heating Fuel by Race and Ethnicity", label="tab:HeatFuel")
@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$Climate_Region_Pub),caption="Climate by Race and Ethnicity", label="tab:Climate")
@

\begin{figure}
\begin{center}
\caption{Age of Structure by Rent/Own and Race/Ethnicity}
\label{fig:AgebyOwnRace}
<<echo=false,fig=TRUE>>=
qplot(YEARMADE,data=RECS,facets =KOWNRENT~EthRace, geom = "density",)
@
\end{center}
\end{figure}

@@

<<echo=false, results=tex>>=
xtable(table(RECS$EthRace,RECS$NUMMEAL),caption="Weekly Meals Cooked by Race and Ethnicity", label="tab:Cooking")
@

\begin{figure}
\begin{center}
\caption{Daytime Temp and Race/Ethnicity (Winter)}
\label{fig:TempHomeRace}
<<echo=false,fig=TRUE>>=
boxplot(TEMPHOME~EthRace, notch=TRUE,data=RECS)
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Daytime Temp When Away and Race/Ethnicity (Winter)}
\label{fig:DayAwayRace}
boxplot(TEMPGONE~EthRace, notch=TRUE,data=RECS)
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Nightime Temp and Race/Ethnicity (Winter)}
\label{fig:NightRace}
<<echo=false,fig=TRUE>>=
boxplot(TEMPNITE~EthRace, notch=TRUE,data=RECS)

@@

\begin{figure}
\begin{center}
\caption{Daytime Temp Home by Race/Ethnicity (Summer)}
\label{fig:NightRaceS}
<<echo=false,fig=TRUE>>=
boxplot(TEMPHOMEAC~EthRace, notch=TRUE,data=RECS)

@@

\begin{figure}
\begin{center}
\caption{Day Temp Away by Race/Ethnicity (Summer)}
\label{fig:NightRaceS}
<<echo=false,fig=TRUE>>=
boxplot(TEMPGONEAC~EthRace, notch=TRUE,data=RECS)

@@

\begin{figure}
\begin{center}
\caption{Nightime Temp and Race/Ethnicity (Summer)}
\label{fig:NightRaceS}
<<echo=false,fig=TRUE>>=
boxplot(TEMPNITEAC~EthRace, notch=TRUE,data=RECS)
@
\end{center}
\end{figure}

@@

  \end{itemize}
<<echo=false,results=tex>>=
library(xtable)
HUD<-read.csv("HUD.csv")
HUDCut<-HUD[c("state","pop","reporttot")]
names(HUDCut)<-c("State","Population","Reports (per 100,000)")

#***TODO Fix so it does not show the row number***
xtable(HUDCut,caption="HUD Complaints per 100,000 Population", label="tab:HUDComplaints")
rm(HUD)
@

@@

Our model is based more on the traditional conditional demand models, and includes terms for electricity use contingent on: the age of the home, which is a proxy for building code requirements and insulation; the weather, in the form of Heating and Cooling degree days with a base temperature of 65F; Meals cooked at home, to indicate energy use related to food prepration; the number and age of refrigerators, a major end-use;  the number of and types of TVs and computers, as a proxy for major plug loads; the existence of a well pump; weather the residence has electric hot water service; the existence of pools and hot tubs as well as if they are electrically heated; the kind of windows installed, as another indication of shell quality; the number of people in the home; and finally the income, race and ethnicity of the residents.  The parameter estimates of the full model can be found in appendix \ref{OrthResults}, partial results are shown in table \ref{tab:ParOrthKWH}.

@@

print(xtable(LimitedKWH,caption="Orthodox kWh Model",label="tab:OrthoKWH"),tabular.environment="longtable",label="tab:ParOrthKWH")
@

@@

As stated in section \ref{sec:WhyHUD}, using HUD complaints as a measure of housing market discrimination based on race and ethnicty is less than optimal measure, but at this early stage is an adequate measure to determine the scale of the effect.

Table \ref{tab:LimParOrthKWH} shows only the discrimiation related results for our model of square footage.  Full results can be seen in appendix \ref{OrthResults}.   Note that the square footage model includes structure type and tenure as an exogenous variable.  This model of square footage will be included later as part of a system estimation of electricity use.

@@

library(stringr)
LimitedSQFT<-coef(summary(OrthodoxSQFT))[ !is.na(str_match(rownames(coef(summary(OrthodoxKWH))),"reporttot")),]

print(xtable(LimitedSQFT,caption="Orthodox kWh Model",label="tab:OrthoKWH"),tabular.environment="longtable",label="tab:LimParOrthKWH")
#Still needs a housman test
@

@@

All parameter estimates for the models are statistically significant at the 1\% level and are displayed in appendix \ref{OrthResults}. 
<<OrthodoxStructTenure,echo=FALSE,results=hide>>=
# Recreating what Edwards did in Stata
#Output is not pretty  ***TODO Make Output pretty ***
# mlogit owndwel rentass totsqft_en lnincome nhsldmem reporttot i.education i.householder_race#c.reporttot urbanrural, base(1)
library(nnet)
OrthodoxStructTenure<-multinom(StrTenure~ RENTHELP+TOTSQFT_EN+Income+NHSLDMEM+ reporttot+EDUCATION+reporttot+reporttot:EthRace+UR,data=RECS,weights=NWEIGHT)

@@

Since all the parameters are statistically significant it is easier to show the effects of our measure of housing market discrimination on the probability of each structure and tenure choice for the race and ethnicity combinations.  The HUD reports run from \Sexp{min(HUDCut[3])} complaints per 100,000 to \Sexp{max(HUDCut[3])} complaints per 100,000.  Figure \ref{} shows the reaction to HUD complaints over the 1 to 21 complaint range.

\begin{figure}
\begin{center}

@@

@
\end{caption}
\end{figure}

@@

The single equation results discussed in section \ref{sec:SingleEq} show strong promis for endogenizing some of the decisions make about structure and equipment within the conditional demand model as well as the

\begin{itemize}
  \item results are bootstrapped 400 times for standard deviation estimates.  More can be done.
  \item probablities of structure type rather than forecast were used.
  \item no censoring for square footage less than zero.
  \item boot strap samples were restricted to samples that allowed for all parameters to be estimated.
\end{itemize}

@@

#Either load or create bootstrap estimates.  Note that multicore does not work on windows machines
library(boot)
BootReplicates<-400
if(file.exists("System.RData")) load("System.RData") else {
  if(Sys.info()[['sysname']]=="Linux") {
    System<-boot(SubsetRECS,BootThing,BootReplicates,parallel="multicore",ncpus=4)} else {
      System<-boot(SubsetRECS,BootThing,BootReplicates)
    }
 save(System,file="System.RData")
}
@

@@

\appendix
\section{Orthodx Results}\label{OrthResults}
\
<<echo=false, results=tex>>=

print(xtable(OrthodoxKWH,caption="Orthodox kWh Model",label="tab:OrthoKWH"),tabular.environment="longtable")

>>>>

%\cite{RBase}

@@

<<tabRaceVStruct,echo=false,results=tex>>=
library(xtable)
#xtable(table(RECS$EthRace,RECS$TYPEHUQ), caption="Count of Observations by Race and Structure Type", label="tab:RaceVStruct")
@

<<echo=false,results=tex>>=
library(xtable)
#xtable(table(RECS$REGIONC,RECS$TYPEHUQ), caption="Count of Observations by Region and Structure Type", label="tab:RegionVStruct")
@

\begin{figure}
\begin{center}
\caption{Annual kWh by Rent/Own and Race}
\label{fig:kWhbyOwnRace}
<<echo=false,fig=TRUE>>=
library(ggplot2)
#qplot(KWH, data = RECS, xlim=c(0,30000),facets =EthRace~KOWNRENT, geom = "density" )
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Annual kWh by Income}
\label{fig:kWhbyIncome}
<<echo=false,fig=TRUE>>=
library(ggplot2)
#qplot(Income,KWH, data = RECS, xlim=c(0,30000),facets =EthRace~KOWNRENT )+geom_boxplot()
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
<<echo=false,fig=TRUE>>=
#qplot(NHSLDMEM,data=RECS,facets =~EthRace, geom = "density",title="Household Size by Race and Ethnicity")+xlab("Household Size")
@
\end{center}
\end{figure}

@@

<<echo=false, results=tex>>=
#xtable(table(RECS$EthRace,RECS$ESWWAC),caption="EnergyStar Wall AC by Race and Ethnicity", label="tab:EstarWall")
@

<<echo=false, results=tex>>=
#xtable(table(RECS$EthRace,RECS$UR),caption="Race by Urban/Rural", label="tab:UrbRural")
@

<<echo=false, results=tex>>=
#xtable(table(RECS$EthRace,RECS$FUELH2O),caption="Water Service by Race and Ethnicity", label="tab:Water")
@

<<echo=false, results=tex>>=
#xtable(table(RECS$EthRace,RECS$FUELHEAT),caption="Heating Fuel by Race and Ethnicity", label="tab:HeatFuel")
@

<<echo=false, results=tex>>=
#xtable(table(RECS$EthRace,RECS$Climate_Region_Pub),caption="Climate by Race and Ethnicity", label="tab:Climate")
@

\begin{figure}
\begin{center}
\caption{Age of Structure by Rent/Own and Race/Ethnicity}
\label{fig:AgebyOwnRace}
<<echo=false,fig=TRUE>>=
#qplot(YEARMADE,data=RECS,facets =KOWNRENT~EthRace, geom = "density",)
@
\end{center}
\end{figure}

@@

<<echo=false, results=tex>>=
#xtable(table(RECS$EthRace,RECS$NUMMEAL),caption="Weekly Meals Cooked by Race and Ethnicity", label="tab:Cooking")
@

\begin{figure}
\begin{center}
\caption{Daytime Temp and Race/Ethnicity (Winter)}
\label{fig:TempHomeRace}
<<echo=false,fig=TRUE>>=
#boxplot(TEMPHOME~EthRace, notch=TRUE,data=RECS)
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Daytime Temp When Away and Race/Ethnicity (Winter)}
\label{fig:DayAwayRace}
#boxplot(TEMPGONE~EthRace, notch=TRUE,data=RECS)
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Nightime Temp and Race/Ethnicity (Winter)}
\label{fig:NightRace}
<<echo=false,fig=TRUE>>=
#boxplot(TEMPNITE~EthRace, notch=TRUE,data=RECS)

@@

\begin{figure}
\begin{center}
\caption{Daytime Temp Home by Race/Ethnicity (Summer)}
\label{fig:HomeRaceS}
<<echo=false,fig=TRUE>>=
#boxplot(TEMPHOMEAC~EthRace, notch=TRUE,data=RECS)

@@

\begin{figure}
\begin{center}
\caption{Day Temp Away by Race/Ethnicity (Summer)}
\label{fig:AwayRaceS}
<<echo=false,fig=TRUE>>=
#boxplot(TEMPGONEAC~EthRace, notch=TRUE,data=RECS)

@@

\begin{figure}
\begin{center}
\caption{Nightime Temp and Race/Ethnicity (Summer)}
\label{fig:NightRaceS}
<<echo=false,fig=TRUE>>=
#boxplot(TEMPNITEAC~EthRace, notch=TRUE,data=RECS)
@
\end{center}
\end{figure}

@@

<<echo=false,results=tex>>=
library(xtable)
HUD<-read.csv("HUD.csv")
HUDCut<-HUD[c("state","pop","reporttot")]
names(HUDCut)<-c("State","Population","Reports (per 100,000)")

#***TODO Fix so it does not show the row number***
#xtable(HUDCut,caption="HUD Complaints per 100,000 Population", label="tab:HUDComplaints")
rm(HUD)
@

@@

Our model is based more on the traditional conditional demand models, and includes terms for electricity use contingent on: the age of the home, which is a proxy for building code requirements and insulation; the weather, in the form of Heating and Cooling degree days with a base temperature of 65F; Meals cooked at home, to indicate energy use related to food prepration; the number and age of refrigerators, a major end-use;  the number of and types of TVs and computers, as a proxy for major plug loads; the existence of a well pump; weather the residence has electric hot water service; the existence of pools and hot tubs as well as if they are electrically heated; the kind of windows installed, as another indication of shell quality; the number of people in the home; and finally the income, race and ethnicity of the residents.  
%Fix latex ref
% The parameter estimates of the full model can be found in appendix \ref{OrthResults}, partial results are shown in table \ref{tab:ParOrthKWH}.

@@

#print(xtable(LimitedKWH,caption="Orthodox kWh Model",label="tab:OrthoKWHLimited"),tabular.environment="longtable",label="tab:ParOrthKWH")
@

@@

As stated in section \ref{sec:WhyHUD}, using HUD complaints as a measure of housing market discrimination based on race and ethnicty is less than optimal measure, but at this early stage is an adequate measure to determine the scale of the effect.  
%Fix ref
% Table \ref{tab:OrthoSQFTLimited} shows only the discrimiation related results for our model of square footage.  Full results can be seen in appendix \ref{OrthResults}.   Note that the square footage model includes structure type and tenure as an exogenous variable.  This model of square footage will be included later as part of a system estimation of electricity use.

@@

library(stringr)
LimitedSQFT<-coef(summary(OrthodoxSQFT))[ !is.na(str_match(rownames(coef(summary(OrthodoxSQFT))),"reporttot")),]

#print(xtable(LimitedSQFT,caption="Orthodox kWh Model",label="tab:OrthoSQFTLimited"),tabular.environment="longtable")
#Still needs a housman test
@

@@

%fix ref
% All parameter estimates for the models are statistically significant at the 1\% level and are displayed in appendix \ref{OrthResults}.

<<OrthodoxStructTenure,echo=FALSE,results=hide>>=
# Recreating what Edwards did in Stata
#Output is not pretty  ***TODO Make Output pretty ***
# mlogit owndwel rentass totsqft_en lnincome nhsldmem reporttot i.education i.householder_race#c.reporttot urbanrural, base(1)
library(nnet)
OrthodoxStructTenure<-multinom(StrTenure~ RENTHELP+TOTSQFT_EN+Income+NHSLDMEM+ reporttot+EDUCATION+reporttot+reporttot:EthRace+UR,data=RECS,weights=NWEIGHT)

@@

%Fix the Latex part of this.
% Since all the parameters are statistically significant it is easier to show the effects of our measure of housing market discrimination on the probability of each structure and tenure choice for the race and ethnicity combinations.  The HUD reports run from \Sexp{min(HUDCut[3])} complaints per 100,000 to \Sexp{max(HUDCut[3])} complaints per 100,000.  Figure \ref{fig:HUDOnChoice} shows the reaction to HUD complaints over the 1 to 21 complaint range.

\begin{figure}
\begin{center}\label{fig:HUDOnChoice}

@@

@@

The single equation results discussed in section \ref{sec:SingleEq} show strong promis for endogenizing some of the decisions made about structure and equipment within the conditional demand model as well as the potential importance of houseing market discrimination in structure choice.  In this section we treat square footage, tenure and structure type as endogenous and estimate estimate the electricty, structure/tenure, and square footage model as a system.

It is unclear how this can be acomplished in a full information, so we chose a two-state least squares technique estimating first the square footage model, then useing forecasted values to estimate tenure and structure type model.  Forecasts from that model were used to resestimate and produce a new round of forecasts for the square footage model.  Both the forecasts of the structure tenure model and square footage model were then used to estimate the conditional demand model.

This should produce consistent results but the variance of the paremeter estimates are biased.  It is unclear how to make the usual corrections to the variance of the paremeter estimates given that the structure and tenure model is estimated as a multinomial logit.  The best alternative is to bootstrap the system.  There are a few caveates.

First, the bootstrap sampling is stratified so that all paremater estimates in all models can be estimated.  This is particularly important in the structre and tenure model.  We had to ensure enough observations of the rare cells, rented mobile homes being the most restricted.  This was acomplished by simple rejection sampling rather than assigning different probabilities of selection to each of the cells.

Second, we did not censor non-possive square footage forecasts.  The square footage model is weaker than expected and it was quite possible to have negative forecasts.  Alternative models, such as Tobit, would not be effective, but we are considering transformations of square footage as we move foreward.

Finally, only 400 bootstrap replicates were evaluated.  This is usually enough to produce adequate estimates of the standard deviations of the paremeter estimates in the electricity model but insufficent for BCa, percentile, or even basic bootstrap confidence intervals.

% \begin{itemize}
%   \item results are bootstrapped 400 times for standard deviation estimates.  More can be done.
%   \item probablities of structure type rather than forecast were used.
%   \item no censoring for square footage less than zero.
%   \item boot strap samples were restricted to samples that allowed for all parameters to be estimated.
% \end{itemize}

@@

#Either load or create bootstrap estimates.  Note that multicore does not work on windows machines
library(boot)
library(parallel)
BootReplicates<-400
if(file.exists("System.RData")) load("System.RData") else {
  if(Sys.info()[['sysname']]=="Linux") {
    System<-boot(SubsetRECS,BootThing,BootReplicates,parallel="multicore",ncpus=detectCores())} else {
      System<-boot(SubsetRECS,BootThing,BootReplicates)
    }
 save(System,file="System.RData")
}
@

***NEED TEXT TO EXPLAIN HERE***

@@

\appendix
\section{Orthodox Results}\label{OrthResults}

<<echo=false, results=tex>>=

print(xtable(OrthodoxKWH,caption="Orthodox kWh Model",label="tab:OrthoKWHFull"),tabular.environment="longtable")
